### 2018-04-09 paper/introduction.tex
basically done, need a little more work in overview and its graph

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that strictly
contains the union of the classes of body-disjoint and linear systems.
%
In particular, the class consists of all CHC systems such that, for
each clause $\mathcal{C}$, the dependences of all distinct relational
predicates in $\mathcal{C}$ are disjoint;
%
we thus refer to such systems as \emph{Clause-Dependent Disjoint
  (CDD)} systems.
%
We present a symbolic algorithm that demonstrates that CDD is directly solvable.
%

% a new solver
The second contribution of this paper is a novel solver for
CHC systems, named \sys.
%
Given recursion-free system $\mathcal{S}$, \sys reduces the problem
of solving $\mathcal{S}$ to solving an equivalent CDD system
$\mathcal{S}'$.
%
In the worst case, the size of $\mathcal{S}'$ may be exponential in the size of
$\mathcal{S}$.
%
In our experience, the size of $\mathcal{S}'$ is usually close enough
to the size of $\mathcal{S}$ for \sys to perform significantly better
than the best known CHC solvers.
%
Because CDD strictly contains the union of the classes 
of body-disjoint and linear systems, the size of an equivalent minimum 
CDD system is equal or less then the size of equivalent body-disjoint and
linear system.
%
The procedure implemented in \sys to construct $\mathcal{S}'$ is a
generalization of existing procedures for generating compact
verification conditions of hierarchical
programs~\cite{flanagan01,lal-qadeer15}.
%
Given a recursion system, \sys solves a series recursion-free systems 
from bounded unwinding the original system and combines the solutions
of recursion-free systems to generate the solution for the recursion system
as previous work proposed~\cite{rummer13b}.
%

>>>>

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that strictly
contains the union of the classes of body-disjoint and linear systems.
%
In particular, the class consists of all recursion-free CHC systems such that, 
for each clause $C$, the dependences of all distinct relational
predicates in the body of $C$ are disjoint;
%
we thus refer to such systems as \emph{Clause-Dependent Disjoint
  (CDD)} systems.
%
We present a symbolic algorithm that demonstrates that CDD is directly solvable.
%

% a new solver
The second contribution of this paper is a novel solver for
CHC systems, named \sys.
%
Given recursion-free system $S$, \sys reduces the problem
of solving $S$ to solving an equivalent CDD system
$S'$.
%
In the worst case, the size of $S'$ may be exponential in the size of
$S$.
%
In our experience, the size of $S'$ is usually close enough
to the size of $S$ for \sys to perform significantly better
than the best known CHC solvers.
%
Because CDD strictly contains the union of the classes 
of body-disjoint and linear systems, the size of an equivalent minimum 
CDD system is equal or less then the size of equivalent body-disjoint and
linear system.
%
The procedure implemented in \sys to construct $S'$ is a
generalization of existing procedures for generating compact
verification conditions of hierarchical
programs~\cite{flanagan01,lal-qadeer15}.
%
Given a recursion system, \sys solves a series recursion-free systems 
from bounded unwinding the original system and combines the solutions
of recursion-free systems to generate the solution for the recursion system
as previous work proposed~\cite{rummer13b}.
%

### 2018-04-09 paper/overview.tex
basically done, need a little more work in overview and its graph

%first section
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$, that formulates verifying the safety of the program
\cc{multiCall} (\autoref{fig:multicall-code}), and illustrate that
$\mcchc$ is not in the directly solvable classes introduced in
previous work.
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
\emph{Clause-Dependent Disjoint (CDD)} system and how \sys solves it.

@@

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.26\textwidth] %
    { \caption{\cc{multiCall}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    { \input{code/multiCall.java} }
    %
    \ffigbox[.7\textwidth] %
    {\caption{Dependencies of CHC system $\mcchc$, depicted as a
        hypergraph.
        % 
        Each relational predicate is a node (circle) and each clause
        is a hyperedge (box). } %
      \label{fig:ex-deps} }
    { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{multiCall} includes a procedure \cc{dbl} that, given an integer
variable \cc{n}, returns the value of \cc{n} doubled (line 1).
%
\cc{main}, the entry procedure of \cc{multiCall}, given integer \cc{n}
(line 3), first binds variable \cc{res} to $0$ (line 4).
%
\cc{main} then tests if \cc{n} is greater than or equal to $0$ (line
5).
%
If so, \cc{main} doubles the value in \cc{n} and binds the result to
\cc{res} (line 6).
%
If not, \cc{main} doubles the value in \cc{n}, negates the result, and
binds the final result to \cc{res} (line 8).
%
\cc{main} then returns the value stored in \cc{res} (line 9).

% state the safety problem:
The problem that we address in this work, in the context of
\cc{multiCall}, is to verify that for each input, \cc{multiCall}
returns a value greater than or equal to $0$.
% talk about solving Horn Clauses:
Verifying that \cc{multiCall} satisfies such a property can be reduced
to solving a recursion-free CHC system over a set of relational
predicates $\mcpreds$ that contains each control location and
procedure name in \cc{multiCall}.
%
A solution interprets each control location \cc{L} as a summary that
relates the arguments in each active call of the function containing
\cc{L} to the program state at \cc{L}, and interprets each procedure
name \cc{f} as a summary of \cc{f}.
%
For each step of \cc{multiCall} to \cc{L}, there is a corresponding
clause with head predicate \cc{L}.
%
For each procedure \cc{f}, there is a clause relating state at return
locations of \cc{multiCall} to a summary of their procedure.
%
In particular, one such system $\mcchc$ is
%
\begin{gather}
  % semantic constraint of dbl:
  \cc{dbl}(\cc{n}, 2 \cdot \cc{n}) \qquad
  % semantic constraint of line 3:
  \cc{L3}(\cc{n}) \implies 
  % semantic constraint of line 4:
  \cc{L4}(\cc{n}, 0) \notag \\
  % semantic constraint of line 5 (then branch)
  \cc{L4}(\cc{n}, \cc{res}), \cc{dbl}(\cc{n}, \cc{res}'), %
  \cc{n} \geq 0 \implies %
  \cc{L9}(\cc{n}, \cc{res}') \label{eqn:then} \\
  % semantic constraint of line 6 (else branch)
  \cc{L4}(\cc{n}, \cc{res}), \cc{dbl}(\cc{n}, \cc{res}'), %
  \cc{n} < 0 \implies %
  \cc{L9}(\cc{n}, -\cc{res}') \label{eqn:else} \\
  % semantic constraint of main procedure:
  \cc{L9}(\cc{n}, \cc{res}) \implies %
  \cc{main}(\cc{n}, \cc{res}) \qquad
  % errors:
  \cc{main}(\cc{n}, \cc{res}), \cc{res} < 0 \implies \mathsf{Err}()
  \notag
\end{gather}
%
While $\mcchc$ has been presented as the result
of a translation from hierarchical program \cc{multiCall}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

% walk through the zoo of CHC systems:
Competitive solvers of recursion-free CHC systems reduce the problem
of solving a given CHC system $\mathcal{S}$ to solving multiple
systems, constructed as copies of sub-systems of $\mathcal{S}$, each
of which is directly solvable (as described in \autoref{sec:intro}).
%
In particular, CHC solvers based on tree interpolation reduce solving
$\mathcal{S}$ to solving copies of subsystems of $\mathcal{S}$ in
which each relational predicate in $\mathcal{S}$ is the head of at
most one clause~\cite{bjorner13,heizmann10}.
%
Such systems do not include $\mcchc$, which contains multiple clauses
with the same head.

@@

% solving the running example
\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
\label{sec:solve-ex}
% restate defn of CDD systems:
\sys solves $\mcchc$ directly by issuing a series of
interpolation queries, each of which generates an interpretation for a
relational predicate in $\mcchc$.
%
The key observation behind the design of \sys is that for CHC system
$\mathcal{S}$, if the dependences of each relational predicate in
each clause of $\mathcal{S}$ are disjoint (in which we case we say
that $\mathcal{S}$ is \emph{Clause-Dependent-Disjoint (CDD)}), then
$\mathcal{S}$ is directly solvable.
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.

% example system is CDD and can be easily solved:
E.g., $\mcchc$ is CDD because for each clause $\mathcal{C}$ of
$\mcchc$, the dependences of all distinct relational predicates in the
body of $\mathcal{C}$ are disjoint.
%
In particular, for clauses \autoref{eqn:then} and \autoref{eqn:else},
the dependences of relational predicates \cc{L4} and \cc{dbl} are
disjoint.
%
The dependences of \cc{L4} is \{\cc{L3}\}, and the dependneces of \cc{dbl}
is $\varnothing$.
%
\autoref{fig:ex-deps} contains a hypergraph $H_{\cc{MC}}$ that depicts
the dependences induced by the clauses of $\mcchc$.
%
The fact that $\mcchc$ is CDD corresponds to the fact that no node in
$H_{\cc{MC}}$ reaches distinct sources of a common hyperedge.

% walk through how a CDD system is solved directly:
A solution of an arbitrary CDD system $\mathcal{S}$ can be constructed
from the results of a series of interpolation queries, each one issued
to find an interpretation of a relational predicate of $\mathcal{S}$
in the topological order of dependences between relational predicates.
%
For relational predicate $R$ in $\mathcal{S}$, the interpretation of
$R$ is the interpolant of two formulas: %
\textbf{(1)} a Boolean combination of the interpretations found for
each dependence of $R$ in $\mathcal{S}$ and constraints in
$\mathcal{S}$ and %
\textbf{(2)} a Boolean combination of constraints that occur in
$\mathcal{S}$.

>>>>

%first introduce the question
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$, that formulates verifying the safety of the program
\cc{multiCall} (\autoref{fig:multicall-code}).
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
\emph{Clause-Dependent Disjoint (CDD)} system and how \sys can solve it by 
encoding it into binary interpolants.
%
In \autoref{sec:not-in}, we illustrate that
$\mcchc$ is not in the directly solvable classes introduced in
previous work.
%

@@

\QZ{remake the graph}
% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.26\textwidth] %
    { \caption{\cc{multiCall}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    { \input{code/multiCall.java} }
    %
    \ffigbox[.7\textwidth] %
    {\caption{Dependencies of CHC system $\mcchc$, depicted as a
        hypergraph.
        % 
        Each relational predicate is a node (circle) and each clause
        is a hyperedge (box). } %
      \label{fig:ex-deps} }
    { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{multiCall} includes a procedure \cc{dbl} that, given an integer
variable \cc{n}, returns the value of \cc{n} doubled (line 1).
%
\cc{main}, the entry procedure of \cc{multiCall}, given integer \cc{n}
(line 3), first binds variable \cc{res} to $0$ (line 4).
%
\cc{main} then tests if \cc{n} is greater than or equal to $0$ (line
5).
%
If so, \cc{main} doubles the value in \cc{n} and binds the result to
\cc{res} (line 6).
%
If not, \cc{main} doubles the value in \cc{n}, negates the result, and
binds the final result to \cc{res} (line 8).
%
\cc{main} then returns the value stored in \cc{res} (line 9).

% state the safety problem:
The problem that we address in this work, in the context of
\cc{multiCall}, is to verify that for each input, \cc{multiCall}
returns a value greater than or equal to $0$.
% talk about solving Horn Clauses:
Verifying that \cc{multiCall} satisfies such a property can be reduced
to solving a recursion-free CHC system over a set of uninterpreted predicates
$\mcpreds$ that each represents a control location and
procedure name in \cc{multiCall}.
%
A solution interprets each control location \cc{L} as a summary that
relates the arguments in each active call of the function containing
\cc{L} to the program state at \cc{L}, and interprets each procedure
name \cc{f} as a summary of \cc{f}.
%
For each step of \cc{multiCall} to \cc{L}, there is a corresponding
clause with head predicate \cc{L}.
%
For each procedure \cc{f}, there is a clause relating state at return
locations of \cc{multiCall} to a summary of their procedure.
%
In particular, one such system $\mcchc$ is
%
\QZ{left allign}
\begin{gather}
  % semantic constraint of dbl:
  \cc{dbl}(\cc{n}, \cc{res}) \gets \cc{res} = 2 * \cc{n} \\
  % semantic constraint of line 3:
   \cc{L4}(\cc{n}, \cc{res}) \gets  \cc{L3}(\cc{n}) \land \cc{res} = 0 \\
  % semantic constraint of line 5 (then branch)
   \cc{L9}(\cc{n}, \cc{res'}) \gets \cc{L4}(\cc{n}, \cc{res}) \land \cc{dbl}(\cc{n}, \cc{res}') \land
   \cc{n} \geq 0 \\
  % semantic constraint of line 6 (else branch)
   \cc{L9}(\cc{n}, \cc{res_1}) \gets \cc{L4}(\cc{n}, \cc{res}) \land \cc{dbl}(\cc{n}, \cc{res_0}) \land
   \cc{n} < 0  \land \cc{res_1} = -\cc{res_0} \\
  % semantic constraint of main procedure:
   False \gets \cc{L9}(\cc{n}, \cc{res}) \land \cc{res} < 0 
\end{gather}
%
While $\mcchc$ has been presented as the result
of a translation from hierarchical program \cc{multiCall}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

% solving the running example
\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
\label{sec:solve-ex}
% state it is CDD
The recursion-free CHC system \cc{multiCall} is a \emph{Clause-Dependent-Disjoint (CDD)}) system.
% state the defn of CDD:
\emph{Clause-Dependent-Disjoint (CDD)}) system is an recursion-free CHC system such that
for any puninterpreted predictes in the body of the same clause, their transitive dependencies
are disjoint(i.e,they do not have common predicates).
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.
% give a example
In particular, for clauses $(3)$ and $(5)$, uninterpreted predicates \cc{L4} and \cc{dbl} shows in
the body of same clause.
%
The transitive dependency of \cc{L4} is \{\cc{L3}\}, and the transitive dependency of \cc{dbl}
is $\varnothing$.
%
Their transitive dependencies are disjoint, because \{\cc{L3}\} $\cap \varnothing = \varnothing$.
%
The formal definition of transitive dependency is given in\QZ{reference}.
%
And the other clauses in \cc{multiCall} only has one uninterpreted predicates in the body, so
\cc{multiCall} is a CDD system.
%
The fact that $\mcchc$ is CDD corresponds to the fact that there no statements executed more than
once in one execution path.
%

% restate defn of CDD systems:
\sys solves $\mcchc$ directly by issuing a binary 
interpolation queriey for each uninterpreted predicates in $\mcchc$ in a topological order.
% walk through how a CDD system is solved directly:
A solution of an arbitrary CDD system $\mathcal{S}$ can be constructed
from the results of a series of interpolation queries, each one issued
to find an interpretation of a relational predicate of $\mathcal{S}$
in the topological order of dependences between relational predicates.
%
For relational predicate $R$ in $\mathcal{S}$, the interpretation of
$R$ is the interpolant of two formulas: %
\textbf{(1)} a Boolean combination of the interpretations found for
each dependence of $R$ in $\mathcal{S}$ and constraints in
$\mathcal{S}$ and %
\textbf{(2)} a Boolean combination of constraints that occur in
$\mathcal{S}$.

\subsection{\cc{multiCall} is not in other system}
\label{sec:not-in}
% walk through the zoo of CHC systems:
Competitive solvers of recursion-free CHC systems reduce the problem
of solving a given CHC system $\mathcal{S}$ to solving multiple
systems, constructed as copies of sub-systems of $\mathcal{S}$, each
of which is directly solvable (as described in \autoref{sec:intro}).
%
In particular, CHC solvers based on tree interpolation reduce solving
$\mathcal{S}$ to solving copies of subsystems of $\mathcal{S}$ in
which each relational predicate in $\mathcal{S}$ is the head of at
most one clause~\cite{bjorner13,heizmann10}.
%
Such systems do not include $\mcchc$, which contains multiple clauses
with the same head.

@@
